---
title: "02_DataManipulation"
output: github_document
date: "2026-09-24"
---

This file is for data manipulation

```{r echo=FALSE, message=FALSE}
library(tidyverse)
```
Importing Necesarry libraries
Importing neccesary datasets, litters and pups datasets

```{r}
litters_df = 
  read_csv(
    "data_import_examples/FAS_litters.csv", 
    na = c("", "NA", ".")
    )

litters_df = janitor::clean_names(litters_df)

pups_df = 
  read_csv(
    "data_import_examples/FAS_pups.csv", 
    na = c("", "NA", "."),
    skip=3
    )

pups_df = janitor::clean_names(pups_df)
```


## `select` some variables

```{r}
select(litters_df, group, litter_number)
select(litters_df, group:gd_of_birth)
select(litters_df, -group)
select(litters_df, starts_with("gd"), group)
select(litters_df, GROUP = group, everything())
```
Instead of using select, we can use rename to just rename certain variable names

```{r}
rename(litters_df, GROUP = group)
```

We can also use the relocate function to move columns to the front of the dataset

```{r}
relocate(litters_df, gd0_weight)
```

To select litter number, sex, and pd ears from the pups dataset.
```{r}
select(pups_df, litter_number, sex, pd_ears)
```


## `filter` some rows

By default is a AND operator, you can use the OR | operator to define grouping of or booleans.
```{r}
filter(litters_df, gd_of_birth == 19)
filter(litters_df, pups_born_alive <= 5)
filter(litters_df, pups_born_alive != 6)
filter(litters_df, group %in% c("Con7", "Con8"), pups_born_alive < 7)
drop_na(litters_df, gd0_weight)
```

In the pups dataset, keep only rows where PD walk is less than 11 and sex is 2

```{r}
filter(pups_df, pd_walk < 11, sex == 2)
```

## `Mutate` to change columns

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight, group = str_to_lower(group))
```


Create a variable that subtracts 7 from PD pivot and Create a variable that is the sum of all the PD variables

```{r}
mutate(pups_df, pd_pivot = pd_pivot - 7, sum_pd = pd_eyes,pd_pivot)
```



## `arrange` to help order the rows by values

```{r}
arrange(litters_df, pups_born_alive)
arrange(litters_df, desc(pups_born_alive))
```



explicitly writing out each can result in too many different variable names

```{r}
litters_df_raw = 
    read_csv("./data_import_examples//FAS_litters.csv", na = c("NA", ".", ""))

litters_df_clean_names = janitor::clean_names(litters_df_raw)
litters_df_selected_cols = select(litters_df_clean_names, -pups_survive)
litters_df_with_vars = 
  mutate(
    litters_df_selected_cols, 
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group))
litters_df_with_vars_without_missing = 
  drop_na(litters_df_with_vars, wt_gain)
litters_df_with_vars_without_missing

```
Its even worse to do this one line

```{r}
litters_df_clean = 
  drop_na(
    mutate(
      select(
        janitor::clean_names(
          read_csv("./data_import_examples//FAS_litters.csv", na = c("NA", ".", ""))
          ), 
      -pups_survive
      ),
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)
    ),
  wt_gain
  )
```




Instead, we can do this using piping to keep it clean and clear:

```{r}
litters_df = 
  read_csv("./data_import_examples//FAS_litters.csv", na = c("NA", ".", "")) |> 
  janitor::clean_names() |> 
  select(-pups_survive) |> 
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) |> 
  drop_na(wt_gain)

```

```{r}
pups_clean = 
  read_csv("data_import_examples/FAS_pups.csv", na= c("NA", ".", ""), skip=3) |> 
  janitor::clean_names() |> 
  drop_na() |> 
  select(litter_number, starts_with("pd")) |> 
  mutate(pd_walk_plus_7 = pd_walk + 7)
pups_clean
```


If the first thing isn't a dataframe, we can specify where the implied pipe goes

```{r}
litters_df |> 
  lm(gd18_weight ~ gd0_weight, data= _)
```


## `Select` vs `pull`. 

Select will always give you a dataframe while pull will give you a vector/list of the column, which is not preferable when we are using dataframes.

## `view` function

You can use the view() function to look at your data. Do not include in the final version.
