---
title: "Simple document"
author: Felix Song
output: 
  html_document:
    toc: true
    toc_float: true
---

I'm an R Markdown document! 

# Section 0: Libraries
```{r echo=F, message=F,}
library(tidyverse) #can create code block with cmd + opt + I
```

#Hey this is new
# Section 1

Here's a **code chunk** that samples from 
a _normal distribution_:

```{r}
samp = rnorm(100)
length(samp)
```

# Section 2

I can take the mean of the sample, too!
The mean is `r mean(samp)`.

# Section 3


```{r} 
plot_df =
    tibble(
        x = rnorm(1000, sd = 0.5),
        y = 1 + 2*x + rnorm(1000)
    )
```

# Section 4: A plot

```{r}
ggplot(plot_df, aes(x=x)) + geom_histogram()
ggplot(plot_df, aes(x=x, y=y)) + geom_point()
```

# Section 5

Write a named code chunk that creates a dataframe comprised of: a numeric variable containing a random sample of size 500 from a normal variable with mean 1; a logical vector indicating whether each sampled value is greater than zero; and a numeric vector containing the absolute value of each element. Then, produce a histogram of the absolute value variable just created. Add an inline summary giving the median value rounded to two decimal places. What happens if you set eval = FALSE to the code chunk? What about echo = FALSE?

```{r}
df = 
  tibble(
    num_var = rnorm(500, mean=1),
    log_vec = num_var > 0,
    num_vec = abs(num_var)
  )
#df$log_vec
ggplot(df, aes(x=df$num_vec)) + geom_histogram()
#Using pull(df, num_vec) doesn't access location of variable, just values. $ does access storage.
```


The mediam is `r round(median(df$num_vec), 2)`


# Section 6

This is **bold** or __bold__ and this is _italic_ or *italic*.
This is `code`
This is superscript^2^ and subscript~2~

# Header 1
## Header 2
### Header 3

* Bullet Item
  *indented 
1. Number list 

Table
--------------

First | Second
------|--------
Content | Content

#Section 7

After the previous code chunk, write a bullet list given the mean, median, and standard deviation of the original random sample

* Here is the mean: `r mean(df$num_var)` 
* Here is the median: `r round(median(df$num_var), 2)`
* Here is the stdv: `r sd(df$num_var)`

You can also use Latex here: $x = \frac{\Omega} {123}$











